Asteroid project summary

### 2022-02-03 summaries/21-11/21-11.tex
Updated observation gap

In figure \ref{fig:example-data} we present example data for this asteroid using the two reference asteroid and orbit parameters described in appendix \ref{app:reference-config}. Additionally, one thousand simulations are also run with asteroid parameters $\gamma_0$, $K_{20}$, and $K_{22}$ changed slightly, distributed according to the likelihood defined by equation \ref{eqn:likelihood} with this nominal uncertainty method and $\sigma_\theta = 0.2$. The method for distributing these shape parameters is the same as the method for distributing the MCMC initial walker positions, described in section \ref{sec:fitting-method}. Bands containing 68.3\%, 95.5\%, and 99.7\% of these additional simulations are also plotted in figure \ref{fig:example-data}.

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{nominal-data-sym.pdf}
  \includegraphics[width=0.7\textwidth]{nominal-data-asym.pdf}
  \caption{Angular velocity data simulated for the symmetric (above) and asymmetric (below) reference asteroid densities and the reference orbit (black lines). Also plotted is the deviation of the data for likelihood-distributed changes to the asteroid shape (bands) using the nominal uncertainty method. Bands contain 68.3\%, 95.5\%, and 99.7\% of the 1000 simulations run.}
  \label{fig:example-data}
\end{figure}

@@

To test this, we mask the perigee of the counter by removing data within time $T$ of the perigee, where $T$ ranges from 0 to 1 hour. To prevent lack of precision on $K_{\ell m}$ induced by lower amounts of data for high $T$, we always cut 1 hr$-T$ from the beginning and end of the data set, so that all data sets produced for all $T$ have the same length of data before and after the perigee. We then fit the same asteroid model to the cut data for all $T$ and plot the standard deviations of the parameter posterior PDFs for (1) $\gamma_0$ and $K_{2m}$ as a control, and (2) $K_{3m}$, to test whether they can still be fitted for without data concerning the perigee. We use the asymmetric benchmark asteroid configuration.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{observation-gap.pdf}
  \caption{Observation gap}
  \label{fig:observation-gap}
\end{figure}

>>>>

In figure \ref{fig:example-data} we present example data for this asteroid using the two reference asteroid and orbit parameters described in appendix \ref{app:reference-config}. Additionally, one thousand simulations are also run with asteroid parameters $\gamma_0$, $K_{20}$, and $K_{22}$ changed slightly, distributed according to the posterior distribution obtained by a fit to synthetic data altered according to the nominal uncertainty method (section \ref{sec:uncertainty-models}) and a fit method described in section \ref{sec:fitting-method}.  An observational uncertainty of $\sigma_\theta = 1.0$ was used. However, due to the large amount of data, even this large observational uncertainty does not lead to visible deviations from the true data. So the posterior distribution is widened by a factor $N$, where $N$ is the number of data points in the data set. Bands containing 68.3\%, 95.5\%, and 99.7\% of these additional simulations are plotted in figure \ref{fig:example-data}. This choice of $N$ is made merely for illustrative purposes; the scale of the bands in figure \ref{fig:example-data} have little meaning in an absolute sense, but are meaningful when comparing two bands or two points in one band.

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{nominal-data-sym.pdf}
  \includegraphics[width=0.7\textwidth]{nominal-data-asym.pdf}
  \caption{Angular velocity data simulated for the symmetric (above) and asymmetric (below) reference asteroid densities and the reference orbit (black lines). Also plotted is the deviation of the data for posterior-PDF-distributed perturbations to the asteroid shape (bands) using the nominal uncertainty method. Bands contain 68.3\%, 95.5\%, and 99.7\% of the 1000 simulations run.}
  \label{fig:example-data}
\end{figure}

@@

To test this, we mask the perigee of the counter by removing data within time $T$ of the perigee, where $T$ ranges from 0 to 3 hours. To prevent lack of precision on $K_{\ell m}$ induced by lower amounts of data for high $T$, we always cut 3 hr$-T$ from the beginning and end of the data set, so that all data sets produced for all $T$ have the same length of data before and after the perigee. We then fit the same asteroid model to the cut data for all $T$ and plot the standard deviations of the parameter posterior PDFs for (1) $\gamma_0$ and $K_{2m}$ as a control, and (2) $K_{3m}$, to test whether they can still be fitted for without data concerning the perigee. We use the asymmetric benchmark asteroid configuration.

Figure \ref{fig:observation-gap} shows the best fit $1-\sigma$ and $2-\sigma$ uncertainty in the first order $K_{2m}$ and $\gamma_0$ and higher order $K_{3m}$ parameters as a function of the amount of perigee data $T$ that has been cut from the synthetic data set. Since torque is highest at perigee, we expect that region of the data to contain the most information about $K_{\ell m}$, and therefore that uncertainty should increase monotonically with $T$, which is observed. We also see that this increase halts after about 2 hours of data has been cut for the first order parameters, likely because the majority of the tidal torque occurs within 2 hours of perigee. The exact time of this threshold is dependent on the asteroid orbital parameters.

The higher order parameters show different behavior, increasing in $\sigma$ at smaller $T$ than the first-order parameters, and continuing to increase even for $T>2\siunit{hr}$. Qualitatively, this is expected since the torque is sensitive to the first-order parameters to order $D^{-3}$ and to the second-order parameters to order $D^{-4}$, so that the second-order parameters are expected to be more dominated by data near the center of the orbit than the first-order parameters. It is clear from figure \ref{fig:observation-gap}, however, that not all the higher order parameters increase in uncertainty at the same rate as each other, nor do they increase in uncertainty uniformly with $T$; once $T\sim 1 \siunit{hr}$ is cut out, further cuts to $T$ do not damage precision in $K_{3m}$ as much until $T\gtrsim 2\siunit{hr}$ is reached.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{observation-gap.pdf}
  \caption{Best fit uncertainty in the first (top row) and second (bottom three rows) order parameters as a function of $T$, the amount of perigee observational data cut from the data set. All data within $T$ hours of perigee is cut, and the data set is correspondingly extended before and after the perigee so that its total size is unchanged.}
  \label{fig:observation-gap}
\end{figure}
